### 2023-06-14 data-raw/02-OSM-Amenities.Rmd

---
title: "Network-Preprocessing"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document I collect and preprocess the OSM data with amenities.

# Preliminaries

Load packages:
```{r}
library(dplyr) # A Grammar of Data Manipulation
library(forcats) # Tools for Working with Categorical Variables (Factors)
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3) # R Bindings for H3
library(osmdata) # Import 'OpenStreetMap' Data as Simple Features or Spatial Objects
library(purrr) # Functional Programming Tools
library(sf) # Simple Features for R
library(tidyr) # Tidy Messy Data
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data objects from package {vaccHamilton}. Urban types is the way the City of Hamilton classifies different parts of the city as "Urban", "Suburban", and "Rural":
```{r}
data("urban_types")
#data("data_da_2016")
```

Use only urban and suburban Hamilton:
```{r}
urban_hammer <- urban_types |>
  filter(Type != "Rural")
```

# Obtain amenities

Set the bounding box to the extent of urban/suburban Hamilton:
```{r}
b <- st_bbox(urban_hammer |>
               st_transform(crs = 4326))
```

Initialize the query:
```{r}
amenities <- opq(bbox = b) |>
  add_osm_feature(key = "amenity") |>
  osmdata_sf()
```

The number of amenities retrieved is:
```{r}
nrow(amenities$osm_points)
```

See OSM documentation for relevant amenities: https://wiki.openstreetmap.org/wiki/Key:amenity

```{r}
summary(factor(amenities$osm_points$amenity))
```
Notice the large number of NAs.

Refine search by segmenting the categories of amenities.

Sustenance:
```{r}
sustenance <- opq(bbox = b) |>
  add_osm_features(c("\"amenity\"=\"bar\"",
                     "\"amenity\"=\"biergarten\"",
                     "\"amenity\"=\"cafe\"",
                     "\"amenity\"=\"fast_food\"",
                     "\"amenity\"=\"food_court\"",
                     "\"amenity\"=\"ice_cream\"",
                     "\"amenity\"=\"pub\"",
                     "\"amenity\"=\"restaurant\"")) |>
  osmdata_sf()
```

Check the output:
```{r}
summary(factor(sustenance$osm_points$amenity))
```

Extract the points and add a column with the category "Sustenance":
```{r}
sustenance_points <- sustenance$osm_points |>
  mutate(Category = "Sustenance")
```

Education:
```{r}
education <- opq(bbox = b) |>
  add_osm_features(c("\"amenity\"=\"library\"",
                     "\"amenity\"=\"school\"")) |>
  osmdata_sf()
```

Check the output:
```{r}
summary(factor(education$osm_points$amenity))
```

We don't like this...what is 1,938 NAs in schools and libraries? Libraries is a reasonable number (there are about 22 public libraries, and some are rural). For this reason extract only libraries.

Extract the points, remove the schools and NAs and add a column with the category "Library":
```{r}
library_points <- education$osm_points |>
  filter(amenity == "library") |>
  mutate(Category = "Library")
```

Financial:
```{r}
financial <- opq(bbox = b) |>
  add_osm_features(c("\"amenity\"=\"atm\"",
                     "\"amenity\"=\"bank\"",
                     "\"amenity\"=\"bureau_de_change\"")) |>
  osmdata_sf()
```

Check the output:
```{r}
summary(factor(financial$osm_points$amenity))
```

Possibly 200+ banks and atms not labeled?

Extract the points and add a column with the category "Library":
```{r}
financial_points <- education$osm_points |>
  mutate(Category = "Financial")
```

Health services:
```{r}
healthcare <- opq(bbox = b) |>
  add_osm_features(c("\"amenity\"=\"clinic\"",
                     "\"amenity\"=\"dentist\"",
                     "\"amenity\"=\"doctors\"",
                     "\"amenity\"=\"pharmacy\"")) |>
  osmdata_sf()
```

Check the output:
```{r}
summary(factor(healthcare$osm_points$amenity))
```

Extract the points and add a column with the category "Library":
```{r}
healthcare_points <- healthcare$osm_points |>
  mutate(Category = "Healthcare")
```

Bind all amenity categories as points:
```{r}
amenities_points <- rbind(sustenance_points |>
                            select(osm_id, name, amenity, Category),
                          library_points |>
                            select(osm_id, name, amenity, Category),
                          financial_points |>
                            select(osm_id, name, amenity, Category),
                          healthcare_points |>
                            select(osm_id, name, amenity, Category))
```

Plot:
```{r}
ggplot() +
  geom_sf(data = urban_types,
          aes(fill = Type)) +
  geom_sf(data = amenities_points,
          aes(color = Category))
```

Obtain the coordinates of the points:
```{r}
amenities_coords <- amenities_points |>
  st_coordinates()

amenities_points <- amenities_points |>
  mutate(lon = amenities_coords[,1], lat = amenities_coords[,2])
```

Use the coordinates to obtain the h3 indices at level 13:
```{r}
amenities_points <- amenities_points |>
  mutate(h3_index = geo_to_h3(st_coordinates(geometry),
                              13))
```

Save the data objects:
```{r}
save(sustenance, file = "sustenance.rda", compress = "xz")
save(education, file = "education.rda", compress = "xz")
save(financial, file = "financial.rda", compress = "xz")
save(healthcare, file = "healthcare.rda", compress = "xz")
save(amenities_points, file = "amenities_points.rda", compress = "xz")
```

Append type of urban fabric to amenities:
```{r}
amenities_by_type <- amenities_points |>
  select(Category) |>
  st_transform(crs = st_crs(urban_types)) |>
  st_intersection(urban_types)
```

Summarize number of amenities by category by urban type:
```{r}
amenities_by_type <- amenities_by_type |>
  st_drop_geometry() |>
  group_by(Type) |>
  group_map(~table(.x$Category))

amenities_by_type
```

Notice how the number of amenities is lowest for "Rural" but this is because our search did not include all of rural Hamilton. Still, there is a difference between suburban and urban Hamilton, with urban Hamilton counting more amenties of almost every category. This is despite the difference in areas between the two regions: suburban Hamilton covers considerably more area than urban Hamilton.
```{r}
area_by_type <- urban_hammer |>
  mutate(area = st_area(geometry) |>
           units::set_units(km2))

area_by_type
```

What is the density of amenities in each of these two urban fabrics?
```{r}
suburban_density <- amenities_by_type[[2]]/drop_units(area_by_type$area[1])
urban_density <- amenities_by_type[[3]]/drop_units(area_by_type$area[2])

data.frame(suburban = suburban_density,
           urban = urban_density) |>
  mutate(ratio = urban.Freq/suburban.Freq) |>
  transmute(Category = suburban.Var1, 
            suburban_density = set_units(suburban.Freq, 1/km2), 
            urban_density = set_units(urban.Freq, 1/km2), 
            ratio)
```