### 2023-06-14 data-raw/03-Network-Attributes-by-Walkshed.Rmd

---
title: "Network-Preprocessing"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document I aim to partition the network based on walksheds.

# Preliminaries

Load packages:
```{r}
#library(disk.frame) # Larger-than-RAM Disk-Based Data Manipulation Framework
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(igraph)
#library(h3) # R Bindings for H3
#ibrary(osmdata)
#library(purrr) # Functional Programming Tools
#library(r5r) # Rapid Realistic Routing with 'R5'
library(sf) # Simple Features for R
library(sfnetworks)
library(tidyr)
#library(skimr) # Compact and Flexible Summaries of Data
#library(units)
#library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r}
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
load(file = "walksheds.rda")
```

Extract the edges of the network:
```{r}
hamilton_edges <- hamilton_net$edges
```

Calculate network attributes by walkshed:
```{r}
# Time algorithm
start.time <- Sys.time()

# Initialize data frame
walksheds_net_attributes <- data.frame(GeoUID = walksheds$GeoUID,
                                       n_nodes = numeric(889), 
                                       transitivity = numeric(889),
                                       motifs = numeric(889),
                                       n_edges = numeric(889),
                                       diameter = numeric(889), 
                                       radius = numeric(889),
                                       girth = numeric(889), 
                                       global_efficiency = numeric(889),
                                       edge_density = numeric(889),
                                       edge_connectivity = numeric(889),
                                       mean_distance = numeric(889),
                                       min_cut = numeric(889))

for(i in 1:nrow(walksheds)){
  wsnet <- st_intersection(hamilton_edges, 
                           walksheds[i,]) |> 
    st_cast(to = "LINESTRING")
  wsnet <- as_sfnetwork(wsnet)
  E(wsnet)$weight <- E(wsnet)$length
  walksheds_net_attributes$n_nodes[i] <- gorder(wsnet)
  walksheds_net_attributes$transitivity[i] <- transitivity(wsnet)
  walksheds_net_attributes$motifs[i] <- count_motifs(wsnet)
  walksheds_net_attributes$n_edges[i] <- gsize(wsnet)
  walksheds_net_attributes$diameter[i] <- diameter(wsnet)
  walksheds_net_attributes$radius[i] <- radius(wsnet)
  walksheds_net_attributes$girth[i] <- girth(wsnet)$girth
  walksheds_net_attributes$global_efficiency[i] <- global_efficiency(wsnet)
  walksheds_net_attributes$edge_density[i] <- edge_density(wsnet)
  walksheds_net_attributes$edge_connectivity[i] <- edge_connectivity(wsnet)
  walksheds_net_attributes$mean_distance[i] <- mean_distance(wsnet)
  walksheds_net_attributes$min_cut[i] <- min_cut(wsnet)
  print(i)
}

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Join the type of urban fabric:
```{r}
walksheds_net_attributes <- walksheds_net_attributes |>
  mutate(Type = walksheds$Type) |>
  select(GeoUID, Type, n_nodes:min_cut)
```

Save the data object:
```{r}
save(walksheds_net_attributes, file = "walksheds_net_attributes.rda", compress = "xz")
```

```{r}
walksheds_net_attributes |>
  group_by(Type) |>
  skim()
```

Edge connectivity is not super useful: there is little variation - it is a measure of network vulnerability, minimum number of edges to remove to disconnect two nodes (there is no longer a path between them.)

Transitivity: it is the probability that the neighbors of a node are neighbors between them. In a Manhattan-style grid this number will be low because there will be lots and lots of squares.

Motifs (calculated for size 3): It indicates the presence of particular subgraphs that repeat themselves. In other words, a graph with a large number of motifs will tend to have few unique elements (some elements happen again, again, and again). A graph with many unique elements will have few motifs (repetitive patterns).

Motfis (from igraph help pages)
```{r}
g <- sample_pa(100)
motifs(g, 3)
count_motifs(g, 3)
sample_motifs(g, 3)

```

```{r}
pattern <- graph.full(3)
my.graph <- grg.game(100, 0.2) 
```