### 2023-06-14 presentations/2023-colloque-montreal.qmd
Revisions to colloque presentation after comments from BMR

---
title: |
    | Profiling 
    | the 15-min *pedshed* 
    | from the perspective of 
    | network attributes
author: "**Antonio Páez**, Belén Martín, Josh Arbess, James R. Dunn"
format: revealjs
execute: 
  cache: true
---

```{r load-packages}
#| include: false
#| cache: false

library(broom) # Convert Statistical Objects into Tidy Tibbles
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(ggspatial) # Spatial Data Framework for ggplot2
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(here) # A Simpler Way to Find Your Files
library(patchwork) # The Composer of Plots
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r load-data}
#| include: false

load(file = glue(here(), "/data-raw/amenities_walkshed.rda"))
load(file = glue(here(), "/data-raw/hamilton_net.rda"))
load(file = glue(here(), "/data-raw/urban_hammer_da.rda"))
data("urban_types")
load(file = glue(here(), "/data-raw/walksheds.rda"))
load(file = glue(here(), "/data-raw/walksheds_net_attributes.rda"))
```

```{r clean-net-attributes-data}
#| include: false
d <- walksheds_net_attributes |>
                  drop_na()
```

# The 15-min neighborhood

##

The 15-min neighborhood^[Sometimes mislabeled as the 15-min city.].

- A form of chronourbanism that emphasizes movement at a human scale...
- ...in environments that accommodate a wide range of capabilities throughout the lifespan...
- ...with the aim of improving livability and health.

##

The 15-min neighborhood is powered by:

::: {.r-fit-text style="margin-top: 50px; font-size: 3em; color: black;"}
The 15-min pedshed
:::

A pedestrian shed $\equiv$ a 15-min isochrone.

##

When combined with the environment, the 15-min pedshed is:

1. A normative...
2. ...cumulative opportunities...
3. ...accessibility index.

$$
A_i = \sum_{j \in \text{15-min pedshed}} O_j
$$

Keeping the environment constant, bigger pedsheds result in higher accessibility.

##

The maximum pedshed area^[For a given walking speed] would be in a homogeneous, isotropic surface (i.e., a featureless plain).

![](plain-978359-pxhere.com.jpg){.absolute top=150 left=200 height="450"}

::: {.notes}
image source: https://pxhere.com/en/photo/978359
:::

##

Streetscapes limit where movement can happen, which inevitably leads to smaller pedsheds.

##

And furthermore, not all streetscapes are created equal.

![](suburban-603627-pxhere.com.jpg){.absolute top=200 left=-150 height="400"}

![](urban-995263-pxhere.com.jpg){.absolute top=200 right=-150 height="400"}

::: {.notes}
image source: urban streetscape https://pxhere.com/en/photo/995263
image source: suburban streetscape https://pxhere.com/en/photo/603627
:::

# Research questions

How much do 15-min pedsheds vary in a typical Canadian city (Hamilton)?

What are their characteristics?

# Materials

##

Open source data.

- Street network (OSM).
- Amenities (OSM).
- Dissemination Areas (DAs: Census).
- Urban types (City of Hamilton).

##

$n = 888$ urban and suburban zones in Hamilton (rural Hamilton excluded).

```{r study-area}
#| echo: false

ggplot() +
  geom_sf(data = st_union(urban_types)) +
  geom_sf(data = urban_types |>
            filter(Type != "Rural"),
          aes(fill = Type)) +
  geom_sf(data = urban_hammer_da,
          fill = NA) +
  scale_fill_manual(values = c("Urban" = "#619CFF", "Suburban" = "#F8766D")) +
  theme_void()
```

##

Open source software (`R`).

- Data collection and processing.
- Routing.
- Spatial analysis.
- Network analysis.

##

Open, reproducible research.

[https://github.com/paezha/15-minute-neighborhood](https://github.com/paezha/15-minute-neighborhood)

# pedsheds

##

pedsheds calculated from zonal centroids using [`r5r`](https://ipeagit.github.io/r5r/) assuming a walking speed of 3.6 $km/h$.

```{r example-of-network}
#| echo: false
#| warning: false

max_shed_r <- (set_units(3.6, km/h) |> set_units(m/min)) * set_units(15, min)

sample_da <- data_da_2016 |>
  filter(GeoUID == "35250102")

sample_da_coords <- data.frame(geometry = st_geometry(sample_da |>
                                                        st_centroid() |>
                                                        st_coordinates() |>
                                                        st_point())) |>
  st_as_sf(crs = 4326)

sample_net <- hamilton_net$edges |>
  st_intersection(st_buffer(sample_da_coords, max_shed_r))

ggplot() +
  geom_sf(data = sample_net,
          color = "lightgray") +
  geom_sf(data = sample_da,
          fill = NA,
          linewidth = 1,
          color = "#619CFF") +
  geom_sf(data = sample_da_coords,
          color = "#619CFF",
          size = 3) +
  geom_sf(data = sample_da_coords |>
            st_buffer(dist = 900),
          fill = NA,
          color = "black") +
  labs(caption = "Blue polygon is DA boundary; blue dot is centroid") +
  annotation_scale() +
  theme_void()
```

##

Summary of pedshed areas (max. pedshed area is `r round(pi * max_shed_r^2 |> units::set_units(km^2), 2)` $km^2$).

Notice the large variability within the city and between suburban and urban Hamilton.

```{r ped-shed-area-summary}
#| echo: false

global <- walksheds |>
  mutate(area = st_area(geometry),
         area = units::set_units(area, km^2) |> 
           drop_units()) |>
  st_drop_geometry() |>
  select(-c(GeoUID, Type)) |>
  skim()

by_type <- walksheds |>
  mutate(area = st_area(geometry),
         area = units::set_units(area, km^2) |> 
           drop_units()) |>
  st_drop_geometry() |>
  select(-c(GeoUID)) |>
  group_by(Type) |>
  skim()

rbind(global |> 
        select(numeric.mean:numeric.hist) |> 
        mutate(across(!ends_with("hist"), 
                      ~round(.x, 2))), 
      by_type|> 
        select(numeric.mean:numeric.hist) |> 
        mutate(across(!ends_with("hist"), 
                      ~round(.x, 2)))) |>
  mutate(Type = c("All", "Suburban", "Urban"),
         n = c(d |> nrow(),
               d |> filter(Type == "Suburban") |> nrow(),
               d |> filter(Type == "Urban") |> nrow())) |>
  select(Type, n, starts_with("numeric")) |>
  rename_with(~ str_replace(.x, "numeric.", ""), starts_with("numeric")) |>
  gt() |>
  cols_width(everything() ~ px(100)) |>
  tab_options(table.font.size = 30)
```

# Network attributes

## {.smaller .scrollable}

Global descriptors of the network.

- Transitivity (probability that the friends of my friends are my friends: minimal circuits).
- Motifs (recurrence of graph elements: repetition of patterns).
- Diameter (longest shortest path in network: how distant are the two most distant nodes).
- Radius (shortest shortest path in network: how close are the two closest nodes).
- Girth (length of shortest circle: shortest path for a node to reach itself).
- Global efficiency (harmonic mean of distances between all nodes: how easy it is on average to reach others).
- Edge density (ratio of actual to maximum edges: how connected is the network).
- Edge connectivity (How many edges need to be removed to eliminate all paths between two nodes).
- Mean distance (mean of length of all shortest paths).

##

Profiling the pedsheds based on their network attributes.

```{r train-classification-tree}
#| echo: false

# This is Tree 3 with seed 3 (46125) in 05-Walkshed-Profiles.Rmd
seed3 <- 46125
set.seed(seed3)
tree3 <- evtree(Type ~ ., 
                data = d |> select(-GeoUID))
```

```{r plot-classification-tree}
#| echo: false

# Plot the tree with rounded break labels:

ggparty(tree3) +
  geom_edge() +
  geom_edge_label(aes(label = str_replace_all(breaks_label, 
                                          "0.0207311636201516", 
                                          "0.021"))) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.0,
                     ids = c(1, 2, 5)) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.02,
                     ids = c(3, 4, 6, 7)) +
  geom_node_plot(gglist = list(geom_bar(aes(x = "", 
                                            fill = Type),
                                        position = position_fill()),
                               xlab("Type"),
                               theme_minimal()),
                 # draw only one label for each axis
                 shared_axis_labels = TRUE,
                 # draw line between tree and legend
                 legend_separator = FALSE)
```

##

Typically suburban networks contain many unique patterns (few motifs).

```{r plot-leaf-1}
#| include: false
#| warning: false

set.seed(69717)

leaf_1_plot <- st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(transitivity < 0.0207311636201516,
                             motifs < 378, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf() +
  ggtitle("Transitivity < 0.021, motifs < 378") +
  theme_void()
```

```{r plot-leaf-3}
#| include: false
#| warning: false

set.seed(84451)

leaf_3_plot <- st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(transitivity >= 0.0207311636201516,
                             motifs < 1024, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf() +
  ggtitle("Transitivity >= 0.021, motifs < 1,024") +
  theme_void()
```

```{r display-suburban-networks}
#| echo: false

leaf_1_plot + plot_spacer() + leaf_3_plot
```

##

Typically urban networks contain isomorphic patterns that repeat (large number of motifs).

```{r plot-leaf-2}
#| include: false
#| warning: false

set.seed(44785)

leaf_2_plot <- st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(transitivity < 0.0207311636201516,
                             motifs >= 378, 
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf() +
  ggtitle("Transitivity < 0.021, motifs >= 378") +
  theme_void()
```

```{r plot-leaf-4}
#| include: false
#| warning: false

set.seed(81461)

leaf_4_plot <- st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(transitivity >= 0.0207311636201516,
                             motifs >= 1024, 
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf() +
  ggtitle("Transitivity >= 0.021, motifs >= 1,024") +
  theme_void()
```

```{r display-urban-networks}
#| echo: false

leaf_2_plot + plot_spacer() + leaf_4_plot
```

# Network attributes of pedsheds

##

Re-examining the areas of the pedsheds.

```{r add-classification-to-data}
#| include: false

d <- d |>
  mutate(leaf = case_when(transitivity < 0.0207311636201516 & motifs < 378 ~ "Leaf 1",
                          transitivity < 0.0207311636201516 & motifs >= 378 ~ "Leaf 2",
                          transitivity >= 0.0207311636201516 & motifs < 1024 ~ "Leaf 3",
                          transitivity >= 0.0207311636201516 & motifs >= 1024 ~ "Leaf 4"),
         leaf = factor(leaf))
```

```{r join-area-of-ped-sheds}
#| include: false

d <- d |>
  left_join(walksheds |>
              mutate(area = st_area(geometry),
                     area = units::set_units(area, km^2) |> 
                       drop_units()) |>
              st_drop_geometry(),
            by = c("GeoUID", "Type")) 
```

```{r area-of-ped-sheds-by-classification}
#| echo: false

d |>
  select(area, leaf) |>
  group_by(leaf) |>
  skim() |>
  mutate(across(numeric.mean:numeric.p100, 
                      ~round(.x, 2)),
         Type = c("Typically suburban", "Typically urban", "Typically suburban", "Typically urban")) |>
  select(leaf, Type, starts_with("numeric")) |>
  rename_with(~ str_replace(.x, "numeric.", ""), starts_with("numeric")) |>
  gt() |>
  cols_width(everything() ~ px(100)) |>
  tab_options(table.font.size = 30)
```

##

pedshed area as a function of network attributes.

```{r estimate-model-area}
#| include: false
mdl <- d |>
  mutate(area = log(area),
         transitivity_2 = transitivity^2,
         motifs = motifs/1000,
         motifs_2 = (motifs)^2) |>
  lm(area ~ transitivity + transitivity_2 + motifs + motifs_2, data = _)
```

```{r tidy-results-of-model-area}
#| include: false

mdl_tidy <- mdl |>
  tidy() |>
  mutate(term = c("intercept", "transitivity", "transitivity^2", "motifs", "motifs^2"))
```

```{r plot-effects-transitivity}
#| include: false

b2 <- mdl_tidy |> filter(term == "transitivity") |> pull(estimate)
b3 <- mdl_tidy |> filter(term == "transitivity^2") |> pull(estimate)
f_transitivity <- function(x) exp(b1 + b2 * x +  b3 * x^2)

effect_transitivity <- ggplot() +
  geom_function(fun = f_transitivity,
                xlim = c(0, 0.12)) +
    ylim(c(0.6, 1.6)) +
  ggtitle("Expected area as a function of transitivity") + 
  ylab("area (sq.km)") +
  xlab("transitivity") +
  theme_minimal()
```

```{r plot-effect-motifs}
#| include: false

b1 <- mdl_tidy |> filter(term == "intercept") |> pull(estimate)
b4 <- mdl_tidy |> filter(term == "motifs") |> pull(estimate)
b5 <- mdl_tidy |> filter(term == "motifs^2") |> pull(estimate)
f_motifs <- function(x) exp(b1 + b4 * x +  b5 * x^2)

effect_motifs <- ggplot() +
  geom_function(fun = f_motifs,
                xlim = c(0, 6.824)) +
  ylim(c(0.6, 1.6)) +
  ggtitle("Expected area as a function of motifs (in 1,000s)") + 
  ylab("area (sq.km)") +
  xlab("motifs (1,000s)") +
  theme_minimal()
```

```{r display-model-effects}
#| echo: false

effect_transitivity + effect_motifs
```

# In summary...

## Conclusions

- pedshed area varies greatly within Hamilton.
- The size of pedsheds varies significantly with the attributes of the network.
- The effect is non-linear.
- Results suggest an optimal combination of attributes to maximize pedshed areas.

## Implications

- Typically suburban networks result in relatively small pedsheds.
- Turning suburbs into 15-min neighborhoods may require major transportation infrastructure changes.
- Alternatively, major increases in density.
- Very unlikely that the 15-min city will be achieved, but many 15-min neighborhoods may be possible.

## Future work

- Expand the analysis to include the environment, i.e., the opportunity landscape.
- Try other cities.