### 2023-06-21 data-raw/03-Network-Attributes-by-Walkshed.Rmd
NECTAR presentation

---
title: "Network Attributes by Walkshed"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: html_notebook
---

**NOTE:** Notebook takes about 7 hours to run.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we aim to partition the network based on walksheds.

We tried smoothing the pseudo-nodes, but this process messed up with the geometry of the network which affects some of the network analysis, so we decided not to proceed. Instead, we use the network as is. Intersecting the network mutilates all edges that cross the boundary of the pedshed. To avoid this we keep track of the indices of the edges that intersect the pedshed and then retrieve them from the network.

# Preliminaries

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(igraph)
library(skimr) # Compact and Flexible Summaries of Data
library(sf) # Simple Features for R
library(sfnetworks)
library(tidyr)
```

```{r}
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
load(file = "walksheds_da.rda")
load(file = "walksheds_validation.rda")
```

Extract the edges of the network:
```{r}
hamilton_edges <- hamilton_net$edges
```

Calculate network attributes by walkshed:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

# Initialize data frame
walksheds_da_net_vars <- data.frame(GeoUID = walksheds_da$GeoUID,
                                    n_nodes = numeric(nrow(walksheds_da)), 
                                    transitivity = numeric(nrow(walksheds_da)),
                                    motifs_3 = numeric(nrow(walksheds_da)),
                                    motifs_4 = numeric(nrow(walksheds_da)),
                                    n_edges = numeric(nrow(walksheds_da)),
                                    diameter = numeric(nrow(walksheds_da)), 
                                    radius = numeric(nrow(walksheds_da)),
                                    girth = numeric(nrow(walksheds_da)), 
                                    global_efficiency = numeric(nrow(walksheds_da)),
                                    edge_density = numeric(nrow(walksheds_da)),
                                    edge_connectivity = numeric(nrow(walksheds_da)),
                                    mean_distance = numeric(nrow(walksheds_da)),
                                    min_cut = numeric(nrow(walksheds_da)))

for(i in 1:nrow(walksheds_da)){
  # Find the intersection between the edges and the walksheds. The edges will be truncated, but the indices can be used to retrieve the full edges.
  wsnet <- hamilton_edges |>
    st_filter(
      walksheds_da[i,],
      .predicate = st_intersects)

  # Extract the full edges that were found to intersect the walkshed
  wsnet <- hamilton_edges |> 
    filter(edge_index %in% wsnet$edge_index)

  # Convert to sfnetwork
  wsnet <- as_sfnetwork(wsnet)

  E(wsnet)$weight <- E(wsnet)$length
  walksheds_da_net_vars$n_nodes[i] <- gorder(wsnet)
  walksheds_da_net_vars$transitivity[i] <- transitivity(wsnet)
  walksheds_da_net_vars$motifs_3[i] <- count_motifs(wsnet, size = 3)
  walksheds_da_net_vars$motifs_4[i] <- count_motifs(wsnet, size = 4)
  walksheds_da_net_vars$n_edges[i] <- gsize(wsnet)
  walksheds_da_net_vars$diameter[i] <- diameter(wsnet)
  walksheds_da_net_vars$radius[i] <- radius(wsnet)
  walksheds_da_net_vars$girth[i] <- girth(wsnet)$girth
  walksheds_da_net_vars$global_efficiency[i] <- global_efficiency(wsnet)
  walksheds_da_net_vars$edge_density[i] <- edge_density(wsnet)
  walksheds_da_net_vars$edge_connectivity[i] <- edge_connectivity(wsnet)
  walksheds_da_net_vars$mean_distance[i] <- mean_distance(wsnet)
  walksheds_da_net_vars$min_cut[i] <- min_cut(wsnet)
}

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Join the type of urban fabric:
```{r}
walksheds_da_net_vars <- walksheds_da_net_vars |>
  mutate(Type = walksheds_da$Type) |>
  select(GeoUID, Type, n_nodes:min_cut)
```

Save the data object:
```{r}
save(walksheds_da_net_vars, file = "walksheds_da_net_vars.rda", compress = "xz")
```

Calculate network attributes by walkshed for the validation set:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

# Initialize data frame
walksheds_validation_net_vars <- data.frame(GeoUID = walksheds_validation$GeoUID,
                                            n_nodes = numeric(nrow(walksheds_validation)), 
                                            transitivity = numeric(nrow(walksheds_validation)),
                                            motifs_3 = numeric(nrow(walksheds_validation)),
                                            motifs_4 = numeric(nrow(walksheds_validation)),
                                            n_edges = numeric(nrow(walksheds_validation)),
                                            diameter = numeric(nrow(walksheds_validation)), 
                                            radius = numeric(nrow(walksheds_validation)),
                                            girth = numeric(nrow(walksheds_validation)), 
                                            global_efficiency = numeric(nrow(walksheds_validation)),
                                            edge_density = numeric(nrow(walksheds_validation)),
                                            edge_connectivity = numeric(nrow(walksheds_validation)),
                                            mean_distance = numeric(nrow(walksheds_validation)),
                                            min_cut = numeric(nrow(walksheds_validation)))

for(i in 1:nrow(walksheds_validation)){
  # Find the intersection between the edges and the walksheds. The edges will be truncated, but the indices can be used to retrieve the full edges.
  wsnet <- hamilton_edges |>
    st_filter(
      walksheds_validation[i,],
      .predicate = st_intersects)

  # Extract the full edges that were found to intersect the walkshed
  wsnet <- hamilton_edges |> 
    filter(edge_index %in% wsnet$edge_index)

  # Convert to sfnetwork
  wsnet <- as_sfnetwork(wsnet)

  E(wsnet)$weight <- E(wsnet)$length
  walksheds_validation_net_vars$n_nodes[i] <- gorder(wsnet)
  walksheds_validation_net_vars$transitivity[i] <- transitivity(wsnet)
  walksheds_validation_net_vars$motifs_3[i] <- count_motifs(wsnet, size = 3)
  walksheds_validation_net_vars$motifs_4[i] <- count_motifs(wsnet, size = 4)
  walksheds_validation_net_vars$n_edges[i] <- gsize(wsnet)
  walksheds_validation_net_vars$diameter[i] <- diameter(wsnet)
  walksheds_validation_net_vars$radius[i] <- radius(wsnet)
  walksheds_validation_net_vars$girth[i] <- girth(wsnet)$girth
  walksheds_validation_net_vars$global_efficiency[i] <- global_efficiency(wsnet)
  walksheds_validation_net_vars$edge_density[i] <- edge_density(wsnet)
  walksheds_validation_net_vars$edge_connectivity[i] <- edge_connectivity(wsnet)
  walksheds_validation_net_vars$mean_distance[i] <- mean_distance(wsnet)
  walksheds_validation_net_vars$min_cut[i] <- min_cut(wsnet)
}

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Join the type of urban fabric:
```{r}
walksheds_validation_net_vars <- walksheds_validation_net_vars |>
  mutate(Type = walksheds_validation$Type) |>
  select(GeoUID, Type, n_nodes:min_cut)
```

Save the data object:
```{r}
save(walksheds_validation_net_vars, file = "walksheds_validation_net_vars.rda", compress = "xz")
```

```{r}
walksheds_da_net_vars |>
  group_by(Type) |>
  skim()
```

```{r}
walksheds_validation_net_vars |>
  group_by(Type) |>
  skim()
```

Edge connectivity is not super useful: there is little variation - it is a measure of network vulnerability, minimum number of edges to remove to disconnect two nodes (there is no longer a path between them.)

Transitivity: it is the probability that the neighbors of a node are neighbors between them. In a Manhattan-style grid this number will be low because there will be lots and lots of squares.

Motifs (calculated for size 3 and 4): It indicates the presence of particular subgraphs that repeat themselves. In other words, a graph with a large number of motifs will tend to have few unique elements (some elements happen again, again, and again). A graph with many unique elements will have few motifs (repetitive patterns).

Motfis (from igraph help pages)
```{r}
g <- sample_pa(100)
motifs(g, 3)
count_motifs(g, 3)
sample_motifs(g, 3)

```

```{r}
pattern <- graph.full(3)
my.graph <- grg.game(100, 0.2) 
```

Example of pedshed with network:
```{r}
sample_walkshed <- sample.int(nrow(walksheds_da), 1)

wsnet <- hamilton_edges |>
  st_filter(
    walksheds_da[sample_walkshed,],
    .predicate = st_intersects)

# Extract the full edges that were found to intersect the walkshed
wsnet <- hamilton_edges |> 
  filter(edge_index %in% wsnet$edge_index)

ggplot() +
  geom_sf(data = walksheds_da[sample_walkshed,]) +
  geom_sf(data = wsnet) +
  theme_void()

```