### 2023-06-21 data-raw/04-Accessibility-Calculations.Rmd

---
title: "Accessibility Calculations"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: pdf_document
---

**NOTE:** Notebook takes about 10 minutes to run.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we calculate the amenities at the level of pedshed to get cumulative accessibility scores.

# Preliminaries

Load packages:
```{r warning=FALSE, message=FALSE}
library(dplyr) # A Grammar of Data Manipulation
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(h3)
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(tidyr) # Tidy Messy Data
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

```{r}
#load(file = "hamilton_net.rda")
load(file = "amenities_points.rda")
load(file = "urban_hammer_da.rda")
data(urban_types)
load(file = "walksheds_da.rda")
load(file = "walksheds_validation.rda")
```

Intersect amenities with walksheds to obtain the cumulative opportunities:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

amenities_walksheds_da <- amenities_points |>
  select(amenity, Category) |>
  st_intersection(walksheds_da)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Validation walksheds:
```{r warning=FALSE}
# Time algorithm
start.time <- Sys.time()

amenities_walksheds_validation <- amenities_points |>
  select(amenity, Category) |>
  st_intersection(walksheds_validation)

end.time <- Sys.time()
time.taken <- end.time - start.time
time.taken
```

Save amenities by walkshed:
```{r}
save(amenities_walksheds_da, file = "amenities_walkshed_da.rda", compress = "xz")
save(amenities_walksheds_validation, file = "amenities_walkshed_validation.rda", compress = "xz")
```

Summary of amenity categories by DA and Type:
```{r}
amenities_walksheds_summary <- amenities_walksheds_da |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```

Map accessibility by category of amenity:
```{r}
urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE)) |>
  ggplot() +
  geom_sf(aes(fill = Number_of_Amenities)) +
  geom_sf(data = urban_types |>
            filter(Type != "Rural"),
          aes(color = Type),
          fill = NA) +
  #scale_fill_fermenter(direction = 1, palette = "Greens") +
  facet_wrap(~ Amenity_Category) + 
  theme_void()
```

Find the most 15-minutes DA:
```{r}
amenities_walksheds_summary |>
  filter(Financial == max(Financial) |
         Sustenance == max(Sustenance) |
         Healthcare == max(Healthcare) |
         Library == max(Library))
```

Looks like GeoUID == "35250420" has a really nice combination: max in banking, max in sustenance, pretty high in health, and one library.

How many DAs have zero amenities?
```{r}
zero_amenity_walksheds <- urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  group_by(GeoUID, Type) |>
  summarize(Number_of_Amenities = sum(Number_of_Amenities),
            .groups = "drop") |>
  filter(Number_of_Amenities == 0)
```

```{r}
zero_amenity_walksheds |>
  pull(Type) |>
  table()
```

Check to see if there are any DAs in this list that have both an Urban and a Suburban part:
```{r}
zero_amenity_walksheds |>
  group_by(GeoUID, Type) |>
  summarize(n = n()) |>
  filter(n > 1)
```

Nope, there are none.

What is the area of DAs with zero amenities? What is their population?
```{r}
zero_amenity_walksheds_summary <- zero_amenity_walksheds |>
              st_drop_geometry() |>
  left_join(data_da_2016,
            by = c("GeoUID")) |>
  rename(Type.1 = Type.x,
         Type = Type.y) |>
  group_by(Type.1) |>
  summarize(Population = sum(Population),
            area = sum(`Shape Area`),
            .groups = "drop") |>
  mutate(Prop_Population = Population/sum(data_da_2016$Population), 
         Prop_Tot_Area = area/sum(data_da_2016$`Shape Area`),
         walksheds = "Zero Amenities")

zero_amenity_walksheds_summary
```

Find all minimal 15-minutes DAs, meaning those that have at least one of each amenity:
```{r}
minimal_amenities_walksheds <- amenities_walksheds_summary |>
  filter(Financial >= 1,
         Sustenance >= 1,
         Healthcare >= 1,
         Library >= 1)
```

How many minimal 15 min DAs are urban/suburban:
```{r}
minimal_amenities_walksheds |>
  pull(Type) |>
  table()
```

What is the area of DAs with minimal 15-min walk sheds? What is their population?
```{r}
minimal_amenities_walksheds_summary <- minimal_amenities_walksheds |>
              st_drop_geometry() |>
  left_join(data_da_2016,
            by = c("GeoUID")) |>
  rename(Type.1 = Type.x,
         Type = Type.y) |>
  group_by(Type.1) |>
  summarize(Population = sum(Population),
            area = sum(`Shape Area`),
            .groups = "drop") |>
  mutate(Prop_Population = Population/sum(data_da_2016$Population), 
         Prop_Area = area/sum(data_da_2016$`Shape Area`),
         walksheds = "Minimal Amenities")

minimal_amenities_walksheds_summary
```

Populations by urban fabric type:
```{r}
city_wide_summary <- urban_hammer_da |>
  st_drop_geometry() |>
  group_by(Type.1) |>
  summarize(Population = sum(Population),
            area = sum(Shape.Area),
            .groups = "drop") |>
  mutate(Prop_Population = Population/sum(data_da_2016$Population),
         Prop_Area = area/sum(data_da_2016$`Shape Area`),
         walksheds = "City Wide")

city_wide_summary
```

Proportion of suburban and urban population in zero amenity sheds:
```{r}
zero_amenity_walksheds_summary$Population/city_wide_summary$Population
```

About 50% of suburban population (which represent about 22% of the total population in Hamilton) live in DAs with zero amenities within the 15-min walk shed.

About 15% of urban population (representing about 3.7% of the total population of the city) live in DAs with zero amenities within the 15-min walk shed.

Proportion of suburban and urban population in minimal 15-min sheds:
```{r}
minimal_amenities_walksheds_summary$Population/city_wide_summary$Population
```

Barely 2.3% of suburban residents live in DAs with minimal 15-min walk sheds. They represent approximately 1% of the total population of the city.

More than 15% of urban population live in DAs with minimal 15-min walk sheds, and account for approximately 3.8% of the total population of the city.

Compare amenities by category and urban fabric type:
```{r}
amenities_walksheds_summary |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE)) |>
  group_by(Amenity_Category,
           Type) |>
  skim()
```

Sample two walksheds with at least one amenity, one from urban fabric type and one suburban, and plot:
```{r}
seed <- sample.int(1000000, 1)
set.seed(seed)
seed

# 609054, 144675, 741580

sample_walksheds <- amenities_walksheds_da |>
  group_by(Type) |>
  slice_sample(n = 1)

ggplot()  + 
  # geom_sf(data = urban_types |>
  #               filter(Type != "Rural"),
  #         aes(fill = Type)) +
  geom_sf(data = walksheds_da |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          aes(fill = Type),
          color = "black")+
  geom_sf(data = amenities_walksheds_da |>
            filter((GeoUID == sample_walksheds$GeoUID[1] & Type == sample_walksheds$Type[1]) | 
                     (GeoUID == sample_walksheds$GeoUID[2] & Type == sample_walksheds$Type[2])),
          #aes(color = Category)
          )
```

Distance to amenities:
```{r}
load("ttm_walk_da.rda")
```

```{r}
amenities_by_h3 <- amenities_points |>
  st_drop_geometry() |>
  group_by(h3_index) |>
  summarize(n_amenities = n(),
            .groups = "drop")
```

Join the amenities to the travel time matrix by means of the h3 indices:
```{r}
ttm_walk_to_amenities <- ttm_walk_da |>
  left_join(amenities_by_h3,
            by = "h3_index") |>
  drop_na(n_amenities)
```

Also join the urban fabric type:
```{r}
ttm_walk_to_amenities <- ttm_walk_to_amenities |>
  left_join(urban_hammer_da |>
              st_drop_geometry() |>
              transmute(GeoUID, Type = Type.1),
            by = "GeoUID")
```

Summarize travel times to amenities from DA centroids:
```{r}
junk <- ttm_walk_to_amenities |>
  select(GeoUID, Type, travel_time) |>
  group_by(GeoUID, Type) |>
  summarize(min_travel_time = min(travel_time),
            mean_travel_time = mean(travel_time),
            .groups = "drop")
```

Distribution of mean travel times by urban fabric:
```{r}
junk |>
  ggplot(aes(x = mean_travel_time,
             color = Type)) +
  geom_density()
```

Distribution of min travel times by urban fabric:
```{r}
junk |>
  ggplot(aes(x = min_travel_time,
             color = Type)) +
  geom_density()
```