### 2023-06-21 data-raw/05-Walkshed-Profiles.Rmd

---
title: "Profiling Walksheds"
author: "Antonio Paez"
date: "`r Sys.Date()`"
output: html_notebook
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this document we profile the ped sheds based on their network attributes.

# Preliminaries

Load packages:
```{r warning=FALSE, message=FALSE}
library(broom) # Convert Statistical Objects into Tidy Tibbles
library(dplyr) # A Grammar of Data Manipulation
library(evtree) # Evolutionary Learning of Globally Optimal Trees
library(ggparty) # 'ggplot' Visualizations for the 'partykit' Package
library(ggplot2) # Create Elegant Data Visualisations Using the Grammar of Graphics
library(glue) # Interpreted String Literals
library(gt) # Easily Create Presentation-Ready Display Tables
library(patchwork) # The Composer of Plots
library(tidyr) # Tidy Messy Data
library(tree) # Classification and Regression Trees
library(sf) # Simple Features for R
library(skimr) # Compact and Flexible Summaries of Data
library(SOMbrero) # SOM Bound to Realize Euclidean and Relational Outputs
library(stringr) # Simple, Consistent Wrappers for Common String Operations
library(units) # Measurement Units for R Vectors
library(vaccHamilton) # A Data Package to Estimate Accessibility of Vaccination Sites in Hamilton, ON
```

Load data:
```{r}
load(file = "amenities_walkshed_da.rda")
load(file = "hamilton_net.rda")
load(file = "urban_hammer_da.rda")
data("urban_types")
load(file = "walksheds_da.rda")
load(file = "walksheds_da_net_vars.rda")
```

## Classification tree

The aim is to train a classification tree to profile urban and suburban walksheds based on the attributes of the network. We need to drop 1 observation that has an NA in `transitivity`:
```{r}
d <- walksheds_da_net_vars |>
                  drop_na() |>
  mutate(normalized_motifs_3 = motifs_3/n_edges,
         normalized_motifs_4 = motifs_4/n_edges)
```

We use a technique with evolutionary trees - an alternative to classification trees that is less affecte by the greediness of the algorithm. Since evolutionary trees have some randomness (for the initial values), we can try several to find whether there are patterns. In this case we try four different random seed.

Tree 1 with seed 1 (912571):
```{r}
seed1 <- sample.int(1000000, 1)
#seed1 <- 908544
set.seed(908544)
tree1 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree1)
```

Tree 2 with seed 2 (94766):
```{r}
seed2 <- sample.int(1000000, 1)
#seed2 <- 669402
seed2
set.seed(seed2)
tree2 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree2)
```

Tree 3 with seed 3 (46125):
```{r}
seed3 <- sample.int(1000000, 1)
#seed3 <- 853251
seed3
set.seed(seed3)
tree3 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree3)
```

Tree 4 with seed 4 (556090)
```{r}
seed4 <- sample.int(1000000, 1)
#seed4 <- 707633
seed4
set.seed(seed4)
tree4 <- evtree(Type ~ ., 
                data = d |> select(-c(GeoUID, motifs_3, motifs_4, n_edges, n_nodes)))
plot(tree4)
```

Create a function to estimate the missclassification rate and the evaluation function:
```{r}
mc <- function(obj) 1 - mean(predict(obj) == d$Type)

evalfun <- function(obj) 2 * nrow(d) * mc(obj) + width(obj) * log(nrow(d))

#trees <- list("evtree" = tree1, "rpart" = rp, "ctree" = ct, "rpart2" = rp2, "ctree2" = ct2)
trees <- list("evtree 1" = tree1, "evtree 2" = tree2, "evtree 3" = tree3, "evtree 4" = tree4)

round(sapply(trees, function(obj) c("misclassification" = mc(obj), "evaluation function" = evalfun(obj))), digits = 7)

```

All trees have similar fit.

Plot the trees:
```{r}
tree1_plot <- ggparty(tree1) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

tree2_plot <- ggparty(tree2) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

tree3_plot <- ggparty(tree3) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

tree4_plot <- ggparty(tree4) +
  geom_edge() +
  geom_edge_label(aes(label = breaks_label)) +
  geom_node_label(aes(label = splitvar), ids = "inner") +
  # identical to  geom_node_splitvar() +
  geom_node_label(aes(label = info), ids = "terminal")

(tree1_plot + tree2_plot) / (tree3_plot + tree4_plot)
```

Notice that edge_density and normalized_motifs_3 are the dominant attributes in the classification trees. Trees 1, 3, and 4 have the same general structure, and in fact trees 3 and 4 are identical. Based on this we choose to use `tree3`.

Check the break labels of `tree3` to round:
```{r}
ggparty(tree3)$data$breaks_label
```

```{r add-classification-to-data}
#| include: false

d <- d |>
  mutate(leaf = case_when(edge_density < 0.00591859212548868 & normalized_motifs_3 < 0.86734693877551 ~ "Leaf 1",
                          edge_density < 0.00591859212548868 & normalized_motifs_3 >= 0.86734693877551 ~ "Leaf 2",
                          edge_density >= 0.00591859212548868 & normalized_motifs_3 < 0.967741935483871 ~ "Leaf 3",
                          edge_density >= 0.00591859212548868 & normalized_motifs_3 >= 0.967741935483871 ~ "Leaf 4"),
         leaf = factor(leaf))
```

```{r join-area-of-ped-sheds}
#| include: false

d <- d |>
  left_join(walksheds |>
              mutate(area = st_area(geometry),
                     area = units::set_units(area, km^2) |> 
                       drop_units()) |>
              st_drop_geometry(),
            by = c("GeoUID", "Type")) 
```

```{r area-of-ped-sheds-by-classification}
#| echo: false

d |>
  select(area, leaf) |>
  group_by(leaf) |>
  skim() |>
  mutate(across(numeric.mean:numeric.p100, 
                      ~round(.x, 2)),
         Type = c("Typically suburban", "Typically urban", "Typically suburban", "Typically urban")) |>
  select(leaf, Type, starts_with("numeric")) |>
  rename_with(~ str_replace(.x, "numeric.", ""), starts_with("numeric")) |>
  gt() |>
  cols_width(everything() ~ px(100)) |>
  tab_options(table.font.size = 30)
```

Typically suburban networks contain many unique patterns (few motifs).

0.00591859212548868 & normalized_motifs_3 < 0.86734693877551 ~ "Leaf 1",
                          edge_density < 0.00591859212548868 & normalized_motifs_3 >= 0.86734693877551 ~ "Leaf 2",
                          edge_density >= 0.00591859212548868 & normalized_motifs_3 < 0.967741935483871 ~ "Leaf 3",

```{r plot-leaf-1}
set.seed(29919)

leaf_1_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density < 0.00591859212548868,
                             normalized_motifs_3 < 0.86734693877551, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_1_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_1_plot$edge_index)

#leaf_1_plot <- 
  leaf_1_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs < 0.867") +
  theme_void()
```

```{r plot-leaf-2}
set.seed(84451)

leaf_2_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density >= 0.00591859212548868,
                             normalized_motifs_3 < 0.967741935483871, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_2_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_2_plot$edge_index)

leaf_2_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs >= 0.968") +
  theme_void()
```

```{r display-suburban-networks}
#| echo: false

leaf_1_plot + plot_spacer() + leaf_3_plot
```

Typically urban networks contain isomorphic patterns that repeat (large number of motifs).

```{r plot-leaf-3}
set.seed(92348)

leaf_1_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density < 0.00591859212548868,
                             normalized_motifs_3 >= 0.86734693877551, 
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_1_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_1_plot$edge_index)

#leaf_1_plot <- 
  leaf_1_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs >= 0.867") +
  theme_void()
```

```{r plot-leaf-4}
set.seed(392847)

leaf_4_plot <- st_intersection(hamilton_net$edges,
             walksheds_da |> 
  filter(GeoUID == (walksheds_da_net_vars |> 
                      mutate(normalized_motifs_3 = motifs_3/n_edges) |>
                      filter(edge_density >= 0.00591859212548868,
                             normalized_motifs_3 >= 0.967741935483871, 
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID))))

leaf_4_plot <- hamilton_net$edges |>
  filter(edge_index %in% leaf_4_plot$edge_index)

leaf_4_plot |>
  ggplot() +
  geom_sf() +
  ggtitle("edge density < 0.006, motifs >= 0.968") +
  theme_void()
```

```{r display-urban-networks}
#| echo: false

leaf_2_plot + plot_spacer() + leaf_4_plot
```

## Analysis of pedshed areas as a function of network attributes

```{r estimate-model-area}
#| include: false
mdl <- d |>
  mutate(area = log(area),
         edge_density_2 = edge_density^2,
         normalized_motifs_3_2 = normalized_motifs_3^2) |>
  lm(area ~ edge_density + edge_density_2 + normalized_motifs_3, data = _)
```

```{r tidy-results-of-model-area}
#| include: false

mdl_tidy <- mdl |>
  tidy() |>
  mutate(term = c("intercept", "edge_density", "edge_density^2", "normalized_motifs_3"))
```

```{r plot-effects-transitivity}
#| include: false

b1 <- mdl_tidy |> filter(term == "intercept") |> pull(estimate)
b2 <- mdl_tidy |> filter(term == "edge_density") |> pull(estimate)
b3 <- mdl_tidy |> filter(term == "edge_density^2") |> pull(estimate)
f_transitivity <- function(x) exp(b1 + b2 * x +  b3 * x^2)

effect_transitivity <- ggplot() +
  geom_function(fun = f_transitivity,
                xlim = c(0, 0.4)) +
    ylim(c(0.05, 1.9)) +
  ggtitle("Expected area as a function of transitivity") + 
  ylab("area (sq.km)") +
  xlab("transitivity") +
  theme_minimal()
```

```{r plot-effect-motifs}
#| include: false

b1 <- mdl_tidy |> filter(term == "intercept") |> pull(estimate)
b4 <- mdl_tidy |> filter(term == "normalized_motifs_3") |> pull(estimate)
f_motifs <- function(x) exp(b1 + b4 * x)

effect_motifs <- ggplot() +
  geom_function(fun = f_motifs,
                xlim = c(0.45, 1.1)) +
  ylim(c(0.05, 1.9)) +
  ggtitle("Expected area as a function of normalized motifs (size = 3)") + 
  ylab("area (sq.km)") +
  xlab("normalized motifs (size = 3)") +
  theme_minimal()
```

```{r display-model-effects}
#| echo: false

effect_transitivity + effect_motifs
```

<!--
Check the break labels of `tree3` to round:
```{r}
ggparty(tree3)$data$breaks_label
```

Plot the tree with the rounded break labels:
```{r}
ggparty(tree3) +
  geom_edge() +
  geom_edge_label(aes(label = str_replace(breaks_label, 
                                          "0.0207311636201516", 
                                          "0.021"))) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.0,
                     ids = c(1, 2, 5)) +
  geom_node_splitvar(aes(label = ifelse(kids == 0,
                                        glue(" n = {nodesize}"), 
                                        splitvar)),
                     nudge_x = 0.02,
                     ids = c(3, 4, 6, 7)) +
  geom_node_plot(gglist = list(geom_bar(aes(x = "", 
                                            fill = Type),
                                        position = position_fill()),
                               xlab("Type"),
                               theme_minimal()),
                 # draw only one label for each axis
                 shared_axis_labels = TRUE,
                 # draw line between tree and legend
                 legend_separator = FALSE
                 )
```

Find examples of ped sheds in each of the leafs and plot the network for illustration.

Leaf 1:
```{r}

seed <- sample.int(100000, 1)
set.seed(seed)
seed
#69717

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(transitivity < 0.0207311636201516,
                             motifs < 378, 
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Leaf 2 (urban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#84451, 44785

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(motifs >= 378, 
                             transitivity < 0.0207311636201516,
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Leaf 3 (suburban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#81461, 84451

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(motifs < 1024, 
                             transitivity >= 0.0207311636201516,
                             Type == "Suburban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Leaf 4 (urban):
```{r}
seed <- sample.int(100000, 1)
set.seed(seed)
seed
#52534

st_intersection(hamilton_net$edges,
             walksheds |> 
  filter(GeoUID == (walksheds_net_attributes |> 
                      filter(motifs >= 1024, 
                             transitivity >= 0.0207311636201516,
                             Type == "Urban") |> 
                      slice_sample(n=1) |> 
                      pull(GeoUID)))) |>
  ggplot() +
  geom_sf()
```

Add leaf labels to data:
```{r}
d <- d |>
  mutate(leaf = case_when(transitivity < 0.0207311636201516 & motifs < 378 ~ "Leaf 1",
                          transitivity < 0.0207311636201516 & motifs >= 378 ~ "Leaf 2",
                          transitivity >= 0.0207311636201516 & motifs < 1024 ~ "Leaf 3",
                          transitivity >= 0.0207311636201516 & motifs >= 1024 ~ "Leaf 4"),
         leaf = factor(leaf))
```

```{r}
skim(d)
```

```{r}
d |>
  left_join(walksheds |>
              select(-Type) |>
              mutate(area = st_area(geometry),
                     area = units::set_units(area, km^2) |> 
                       drop_units(),
                     area = log(area)) |>
              st_drop_geometry(),
            by = "GeoUID") |>
  lm(area ~ leaf, data = _) |>
  summary()
```

But what about accessibility??

Summary of amenity categories by DA and Type:
```{r}
amenities_walksheds_summary <- amenities_walksheds |>
  drop_na() |>
  st_drop_geometry() |>
  group_by(GeoUID,
           Type,
           Category) |>
  summarize(n = n(),
            .groups = "drop") |>
  pivot_wider(id_cols = c(GeoUID, Type),
              names_from = Category,
              values_from = n,
              values_fill = 0)
```

To map accessibility by category of amenity we first categorize the number of amenities:
```{r}
amentities_da <-  urban_hammer_da |>
  select(GeoUID, Type.1) |>
  rename(Type = Type.1) |>
  left_join(amenities_walksheds_summary,
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~ replace_na(.x, 0))) |>
  pivot_longer(cols = -c(GeoUID, Type, geometry),
               names_to = "Amenity_Category",
               values_to = "Number_of_Amenities") |>
  #mutate(Number_of_Amenities = ifelse(Number_of_Amenities == 0, NA, Number_of_Amenities)) |>
  mutate(Number_of_Amenities = case_when(Number_of_Amenities == 0 ~ "0",
                                         Number_of_Amenities > 0 & Number_of_Amenities <= 3 ~ "1-3",
                                         Number_of_Amenities > 3 & Number_of_Amenities <= 6 ~ "4-6",
                                         Number_of_Amenities > 6 & Number_of_Amenities <= 9 ~ "7-9",
                                         Number_of_Amenities > 9 ~ "10 or more"),
         Number_of_Amenities = factor(Number_of_Amenities,
                                      levels = c("0", "1-3", "4-6", "7-9", "10 or more"),
                                      ordered = TRUE))
```

Plot:
```{r}
amentities_da |>
  ggplot() +
  geom_sf(aes(fill = Number_of_Amenities)) +
  geom_sf(data = urban_types |>
            filter(Type != "Rural"),
          aes(color = Type),
          fill = NA) +
  #scale_fill_fermenter(direction = 1, palette = "Greens") +
  facet_wrap(~ Amenity_Category) + 
  theme_void()
```

Pivot wider:
```{r}
junk <- amentities_da |>
  pivot_wider(names_from = "Amenity_Category",
              values_from = "Number_of_Amenities")
```

```{r}
junk2 <- d |>
  left_join(junk |>
              #select(-Type) |>
              st_drop_geometry(),
            by = c("GeoUID", "Type")) |>
  mutate(across(Sustenance:Library, ~forcats::fct_drop(.x)))
```

```{r}
summary(junk2 |> select(Sustenance:Library))
```

## Clusters: try self-organizing maps (SOM)

```{r}
wsn.som <- trainSOM(x.data = walksheds_net_attributes |>
                       select(-c(GeoUID, Type)),
                     dimension = c(5,5), 
                     verbose = TRUE, 
                     nb.save = 5, 
                     topo = "hexagonal")
```

```{r}
wsn.som
```

```{r}
plot(wsn.som, what="energy")
```

```{r}
wsn.som$clustering
```

```{r}
wsn.som$clustering |> table()
```

```{r}
plot(wsn.som, what="obs", type="hitmap")
```

```{r}
summary(wsn.som)
```

```{r}
par(mfrow = c(2,2))
plot(wsn.som, what = "obs", type = "color", variable = 1)
plot(wsn.som, what = "obs", type = "color", variable = 2)
```

```{r}
plot(wsn.som, what = "add", type = "pie", variable = walksheds_net_attributes$Type) +
  scale_fill_brewer(type = "qual") + 
  guides(fill = guide_legend(title = "Type"))
```

```{r}
quality(wsn.som)
```

```{r}
plot(superClass(wsn.som))
```

```{r}
wsn.som <- trainSOM(x.data = walksheds_net_attributes |>
                       select(-c(GeoUID, Type)),
                     dimension = c(4, 4), #2,3
                     verbose = TRUE, 
                     nb.save = 5,
                     topo = "square")
```

```{r}
plot(wsn.som, what="energy")
```

```{r}
quality(wsn.som)
```