### 2024-10-18 data-raw/00-test-r5r.qmd
Verified that the project is reproducible after updating renv

---
title: "Test r5r"
author: "AP"
format: html
---

## Test package

This notebook is used to test the package [{r5r}](), which is used for routing and generating isochrones.

## Running Code

It is important to allocate memory before launching {r5r}:
```{r}
# allocate RAM memory to Java
options(java.parameters = "-Xmx2G")
```

1) build transport network, pointing to the path where OSM and GTFS data are stored
```{r}
library(r5r)
path <- system.file("extdata/poa", package = "r5r")
r5r_core <- setup_r5(data_path = path, verbose = FALSE)
```

2) load origin/destination points and set arguments
```{r}
points <- read.csv(system.file("extdata/poa/poa_hexgrid.csv", package = "r5r"))
mode <- c("WALK", "TRANSIT")
max_walk_time <- 30   # minutes
max_trip_duration <- 60 # minutes
departure_datetime <- as.POSIXct("13-05-2019 14:00:00",
                                 format = "%d-%m-%Y %H:%M:%S")
```

3.1) calculate a travel time matrix
```{r}
ttm <- travel_time_matrix(r5r_core = r5r_core,
                          origins = points,
                          destinations = points,
                          mode = mode,
                          departure_datetime = departure_datetime,
                          max_walk_time = max_walk_time,
                          max_trip_duration = max_trip_duration)
```

3.2) or get detailed info on multiple alternative routes
```{r}
det <- detailed_itineraries(r5r_core = r5r_core,
                            origins = points[370, ],
                            destinations = points[200, ],
                            mode = mode,
                            departure_datetime = departure_datetime,
                            max_walk_time = max_walk_time,
                            max_trip_duration = max_trip_duration,
                            shortest_path = FALSE,
                            drop_geometry = FALSE)
```

4) Calculate number of schools accessible within 20 minutes 
```{r}
access <- accessibility(r5r_core = r5r_core,
                        origins = points,
                        destinations = points,
                        opportunities_colname = "schools",
                        decay_function = "step",
                        cutoffs = 21,
                        mode =  c("WALK", "TRANSIT"),
                        verbose = FALSE)
```

5} Calculate isochrones
```{r}
ttm <- travel_time_matrix(r5r_core = r5r_core,
                          origins = points,
                          destinations = points,
                          mode = mode,
                          departure_datetime = departure_datetime,
                          max_walk_time = max_walk_time,
                          max_trip_duration = max_trip_duration)

iso <- isochrone(
  r5r_core = r5r_core,
  origins = points,
  mode = "WALK",
  mode_egress = "WALK",
  cutoffs = c(0, 15),
  sample_size = 0.8,
  departure_datetime = departure_datetime,
  time_window = 10L,
  max_walk_time = 15,
  max_bike_time = 30,
  max_car_time = 120,
  max_trip_duration = 120L,
  walk_speed = 3.6,
  bike_speed = 12,
  max_rides = 3,
  max_lts = 2,
  draws_per_minute = 5L,
  n_threads = Inf,
  verbose = FALSE,
  progress = TRUE
)
```